Meta-Analysis of the `Ironic' Effects of Intergroup Contact

### 2022-05-03 manuscript/manuscript.Rmd
revision

---
title: "Meta-Analysis of the 'Ironic' Effects of Intergroup Contact"
header: "'IRONIC' EFFECTS META-ANALYSIS"
keywords: intergroup contact, meta-analysis, perceived discrimination, collective action, policy support
bibliography: references.bib
csl: apa.csl
nocite: | 
  @19, @45, @46, @91, @93, @244, @284, @286, @303, @322, @324, @325, @336, @401, @414, @423, @480, @609, @703, @733, @783, @789, @803, @813, @823, @829, @830, @856, @893, @894, @933, @956, @1020, @1042, @1045, @1114, @1163, @1221, @1248, @1376, @1386, @1395, @1407, @1411, @1412, @1494, @1525, @1549, @1576, @1614, @1658, @1695, @1743, @1762, @1807, @1850, @1891, @1914, @1949, @1966, @1993, @2001, @2023, @2033, @2075, @2257, @2309, @2333, @2341, @2375, @2380, @2381, @2382, @2383, @2385, @2391, @2392, @2394, @2395, @2396, @2397, @2398, @2399, @3000, @3005, @3051, @3054, @3060, @3134, @3196, @3201, @3205, @4000, @4001, @4002, @4003, @4004, @4005, @bracegirdle_disentangling_2022, @sengupta_protect_unpublished
abstract: 
  Growing evidence suggests that intergroup contact, psychology’s most-researched paradigm for reducing prejudice, has the ‘ironic’ effect of reducing support for social change in disadvantaged groups. We conducted a preregistered meta-analytic test of this effect across `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(dplyr::n_distinct(id)))` studies with `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(length(sample)))` samples of `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(sum(n)))` disadvantaged-group members. As predicted, intergroup contact was, on average, associated with less perceived injustice (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "pi"]), digits = 2), "$"))`), collective action (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ca"]), digits = 2), "$"))`), and support for reparative policies (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ps"]), digits = 2), "$"))`). However, these associations were small, variable, and consistent with alternative explanations. Across outcomes, `r with(readRDS("../results/results_for_abstract.rds"), paste0(round(min(p_pred)*100), "--", round(max(p_pred)*100), "\\%"))` of studies found *positive* associations with intergroup contact. Moderator analyses explained about a third of the between-sample variance, showing that, at least for perceived injustice, associations with intergroup contact were most consistently negative in studies that measured direct, qualitatively positive contact among adults. We also found evidence for an alternative explanation for the apparent ‘ironic’ effects of intergroup contact as, after controlling for the positive association of negative contact with support for social change, positive contact was no longer associated with any of the outcomes. We close by discussing strengths and limitations of the available evidence and by highlighting open questions about the relationship between intergroup contact and support for social change in disadvantaged groups.
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: apa-template.tex
---

@@

We prepared a protocol in accordance with the PRISMA-P statement [@moher_preferred_2015] and preregistered it on the Open Science Framework.^[[https://osf.io/ryuev](https://osf.io/ryuev)] Preregistration precludes undisclosed flexibility in study selection, outcome selection, and other decisions that influence effect size estimates—and thus prevents confirmation bias from determining the conclusions of a meta-analysis [@lakens_reproducibility_2016]. We report deviations from our preregistered protocol at the end of the Method section. We make data, analysis scripts, and the fully reproducible manuscript available online.^[[https://osf.io/w5tqv](https://osf.io/w5tqv)]

@@

***Study Design.*** We categorized each study as either observational and cross-sectional, observational and longitudinal, quasi-experimental, experimental, an intervention, or other ($\kappa = .85$).^[While the term ‘observational study’ is sometimes used to refer to the naturalistic observation of behavior, we use it in its broader sense, that is, to refer to studies that do not involve random assignment into experimental conditions [@cochran_planning_1965; @everitt_observational_2005].] We categorized each sample as either a student convenience sample, a non-student convenience sample, a probability or representative sample, or another kind of sample ($\kappa = .73$). We categorized the age group(s) in each sample as children ($\leq 12$ years), adolescents (13–18 years), or adults ($\geq 18$ years; $\kappa = .64$).

@@

Figure 2 provides a qualitative overview of the literature. Even though most participants came from studies conducted in India (`r glue('$N = {f_comma(with(distinct(dl, id, sample, n, country), sum(n[country == "India"])))}$')`) and the United States (`r glue('$N = {f_comma(with(distinct(dl, id, sample, n, country), sum(n[country == "USA"])))}$')`; Figure 2a), most samples were collected in North America (`r glue('$I = {f_comma(with(distinct(dl, id, sample, n, continent), length(sample[continent == "North America"])))}$')`) and Europe (`r glue('$I = {f_comma(with(distinct(dl, id, sample, n, continent), length(sample[continent == "Europe"])))}$')`), with few samples from South America and Africa (Figure 2b). Most samples focused on relative inequalities that resulted from long-term migration (`r glue('$I = {f_comma(with(distinct(dl, id, sample, study_setting), length(sample[study_setting == "Long-term migration"])))}$')`), slavery (`r glue('$I = {f_comma(with(distinct(dl, id, sample, study_setting), length(sample[study_setting == "Slavery"])))}$')`), colonization (`r glue('$I = {f_comma(with(distinct(dl, id, sample, study_setting), length(sample[study_setting == "Colonization"])))}$')`), and short-term migration (`r glue('$I = {f_comma(with(distinct(dl, id, sample, study_setting), length(sample[study_setting == "Short-term migration"])))}$')`). With few exceptions, studies used observational, cross-sectional survey designs. Samples sizes ranged from `r with(distinct(dl, id, sample, n), paste0(f_comma(min(n)), " to ", f_comma(max(n)), " ($\\textit{Mdn} = ", f_comma(median(n)), "$)"))`. Unlike most psychological research, 2 in 5 samples used probability or representative sampling. Of all samples, `r with(distinct(dl, id, sample, study_intention), f_prop2percent(mean(study_intention == "Yes"), digits = 0))` (`r glue('$I = {f_comma(with(distinct(dl, id, sample, n, study_intention), length(sample[study_intention == "Yes"])))}$')`) were collected with the intention to examine the effects of intergroup contact on the outcomes considered in this meta-analysis. Many studies instead examined acculturation processes [@berry_immigration_1997] and measured intergroup contact and perceived injustice without hypothesizing that the former would affect the latter. Together, these observations highlight both limitations and strengths of the empirical literature.

@@

Third, the available evidence is almost entirely from cross-sectional, observational studies and thus consistent with alternative explanations for the observed associations. Support for social change could reduce intergroup contact rather than the other way around. For example, disadvantaged-group members involved in collective action might avoid forming friendships with advantaged-group members. Alternatively, the observed associations could be spurious with both being caused by an unobserved confounder. For example, a disadvantaged-group member's socioeconomic status might both expose them to more advantaged-group members and reduce their perception of injustice. Future research should prioritize longitudinal studies to confirm the direction of the observed associations and (field) experiments to rule out confounding and other alternative explanations.^[Published research includes only three longitudinal studies [@401; @956; @1743], all of which conflated within-person change and between-person stability [@hamaker_critique_2015], two experimental studies with a no-contact control condition [@2075; @2033], and two intervention studies [@2399; @789].]

@@

Finally, we cannot rule out that biases in the literature caused our meta-analysis to overestimate the 'ironic' effects of intergroup contact. Publication bias could have prevented studies that found positive or non-significant associations from entering the published literature. We used various methods to detect and correct for publication bias, yet did not find conclusive evidence for or against publication bias. That said, 61% of samples were either not collected with the intention to study the effects of intergroup contact on a relevant outcome (45%) or from unpublished studies (32%)—and thus unlikely to have been affected by publication bias. Another bias that could have affected our findings is that researchers made data-dependent decisions while collecting and analyzing data that are known to inflate false-positive findings [@gelman_statistical_2014; @simmons_false-positive_2011]. This risk is pertinent to a literature in which most studies are observational, include many measures, and are not preregistered—and, therefore, do not constrain the direction of observed associations or the selection of included variables. Future research should use preregistration and other methods to prevent undisclosed flexibility in data collection and analysis.

>>>>

---
title: "Meta-Analysis of the 'Ironic' Effects of Intergroup Contact"
header: "'IRONIC' EFFECTS META-ANALYSIS"
keywords: intergroup contact, meta-analysis, perceived discrimination, collective action, policy support
bibliography: references.bib
csl: apa.csl
nocite: | 
  @19, @45, @46, @91, @93, @244, @284, @286, @303, @322, @324, @325, @336, @401, @414, @423, @480, @609, @703, @733, @783, @789, @803, @813, @823, @829, @830, @856, @893, @894, @933, @956, @1020, @1042, @1045, @1114, @1163, @1221, @1248, @1376, @1386, @1395, @1407, @1411, @1412, @1494, @1525, @1549, @1576, @1614, @1658, @1695, @1743, @1762, @1807, @1850, @1891, @1914, @1949, @1966, @1993, @2001, @2023, @2033, @2075, @2257, @2309, @2333, @2341, @2375, @2380, @2381, @2382, @2383, @2385, @2391, @2392, @2394, @2395, @2396, @2397, @2398, @2399, @3000, @3005, @3051, @3054, @3060, @3134, @3196, @3201, @3205, @4000, @4001, @4002, @4003, @4004, @4005, @bracegirdle_disentangling_2022, @sengupta_protect_unpublished, @osf_repository
abstract: 
  Growing evidence suggests that intergroup contact, psychology’s most-researched paradigm for reducing prejudice, has the ‘ironic’ effect of reducing support for social change in disadvantaged groups. We conducted a preregistered meta-analytic test of this effect across `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(dplyr::n_distinct(id)))` studies with `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(length(sample)))` samples of `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(sum(n)))` disadvantaged-group members. As predicted, intergroup contact was, on average, associated with less perceived injustice (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "pi"]), digits = 2), "$"))`), collective action (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ca"]), digits = 2), "$"))`), and support for reparative policies (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ps"]), digits = 2), "$"))`). However, these associations were small, variable, and consistent with alternative explanations. Across outcomes, `r with(readRDS("../results/results_for_abstract.rds"), paste0(round(min(p_pred)*100), "--", round(max(p_pred)*100), "\\%"))` of studies found *positive* associations with intergroup contact. Moderator analyses explained about a third of the between-sample variance, showing that, at least for perceived injustice, associations with intergroup contact were most consistently negative in studies that measured direct, qualitatively positive contact among adults. We also found evidence for an alternative explanation for the apparent ‘ironic’ effects of intergroup contact as, after controlling for the positive association of negative contact with support for social change, positive contact was no longer associated with any of the outcomes. We close by discussing strengths and limitations of the available evidence and by highlighting open questions about the relationship between intergroup contact and support for social change in disadvantaged groups.
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: apa-template.tex
---

@@

We prepared a protocol in accordance with the PRISMA-P statement [@moher_preferred_2015] and preregistered it on the Open Science Framework.^[[https://osf.io/ryuev/](https://osf.io/ryuev/)] Preregistration precludes undisclosed flexibility in study selection, outcome selection, and other decisions that influence effect size estimates—and thus prevents confirmation bias from determining the conclusions of a meta-analysis [@lakens_reproducibility_2016]. We report deviations from our preregistered protocol at the end of the Method section. We make data, analysis scripts, and the fully reproducible manuscript available online.^[[https://osf.io/w5tqv/](https://osf.io/w5tqv/)]

@@

***Study Design.*** We categorized each study as either correlational and cross-sectional, correlational and longitudinal, quasi-experimental, experimental, an intervention, or other ($\kappa = .85$). We categorized each sample as either a student convenience sample, a non-student convenience sample, a probability or representative sample, or another kind of sample ($\kappa = .73$). We categorized the age group(s) in each sample as children ($\leq 12$ years), adolescents (13–18 years), or adults ($\geq 18$ years; $\kappa = .64$).

@@

Figure 2 provides a qualitative overview of the literature. Even though most participants came from studies conducted in India (`r glue('$N = {f_comma(with(distinct(dl, id, sample, n, country), sum(n[country == "India"])))}$')`) and the United States (`r glue('$N = {f_comma(with(distinct(dl, id, sample, n, country), sum(n[country == "USA"])))}$')`; Figure 2a), most samples were collected in North America (`r glue('$I = {f_comma(with(distinct(dl, id, sample, n, continent), length(sample[continent == "North America"])))}$')`) and Europe (`r glue('$I = {f_comma(with(distinct(dl, id, sample, n, continent), length(sample[continent == "Europe"])))}$')`), with few samples from South America and Africa (Figure 2b). Most samples focused on relative inequalities that resulted from long-term migration (`r glue('$I = {f_comma(with(distinct(dl, id, sample, study_setting), length(sample[study_setting == "Long-term migration"])))}$')`), slavery (`r glue('$I = {f_comma(with(distinct(dl, id, sample, study_setting), length(sample[study_setting == "Slavery"])))}$')`), colonization (`r glue('$I = {f_comma(with(distinct(dl, id, sample, study_setting), length(sample[study_setting == "Colonization"])))}$')`), and short-term migration (`r glue('$I = {f_comma(with(distinct(dl, id, sample, study_setting), length(sample[study_setting == "Short-term migration"])))}$')`). With few exceptions, studies used correlational, cross-sectional survey designs. Samples sizes ranged from `r with(distinct(dl, id, sample, n), paste0(f_comma(min(n)), " to ", f_comma(max(n)), " ($\\textit{Mdn} = ", f_comma(median(n)), "$)"))`. Unlike most psychological research, 2 in 5 samples used probability or representative sampling. Of all samples, `r with(distinct(dl, id, sample, study_intention), f_prop2percent(mean(study_intention == "Yes"), digits = 0))` (`r glue('$I = {f_comma(with(distinct(dl, id, sample, n, study_intention), length(sample[study_intention == "Yes"])))}$')`) were collected with the intention to examine the effects of intergroup contact on the outcomes considered in this meta-analysis. Many studies instead examined acculturation processes [@berry_immigration_1997] and measured intergroup contact and perceived injustice without hypothesizing that the former would affect the latter. Together, these observations highlight both limitations and strengths of the empirical literature.

@@

Third, the available evidence is almost entirely from cross-sectional, correlational studies and thus consistent with alternative explanations for the observed associations. Support for social change could reduce intergroup contact rather than the other way around. For example, disadvantaged-group members involved in collective action might avoid forming friendships with advantaged-group members. Alternatively, the observed associations could be spurious with both being caused by an unobserved confounder. For example, a disadvantaged-group member's socioeconomic status might both expose them to more advantaged-group members and reduce their perception of injustice. Future research should prioritize longitudinal studies to confirm the direction of the observed associations and (field) experiments to rule out confounding and other alternative explanations.^[Published research includes only three longitudinal studies [@401; @956; @1743], all of which conflated within-person change and between-person stability [@hamaker_critique_2015], two experimental studies with a no-contact control condition [@2075; @2033], and two intervention studies [@2399; @789].]

@@

Finally, we cannot rule out that biases in the literature caused our meta-analysis to overestimate the 'ironic' effects of intergroup contact. Publication bias could have prevented studies that found positive or non-significant associations from entering the published literature. We used various methods to detect and correct for publication bias, yet did not find conclusive evidence for or against publication bias. That said, 61% of samples were either not collected with the intention to study the effects of intergroup contact on a relevant outcome (45%) or from unpublished studies (32%)—and thus unlikely to have been affected by publication bias. Another bias that could have affected our findings is that researchers made data-dependent decisions while collecting and analyzing data that are known to inflate false-positive findings [@gelman_statistical_2014; @simmons_false-positive_2011]. This risk is pertinent to a literature in which most studies are correlational, include many measures, and are not preregistered—and, therefore, do not constrain the direction of observed associations or the selection of included variables. Future research should use preregistration and other methods to prevent undisclosed flexibility in data collection and analysis.

### 2022-05-03 manuscript/manuscript.tex
revision

We prepared a protocol in accordance with the PRISMA-P statement (Moher
et al., 2015) and preregistered it on the Open Science
Framework.\footnote{\url{https://osf.io/ryuev}} Preregistration
precludes undisclosed flexibility in study selection, outcome selection,
and other decisions that influence effect size estimates---and thus
prevents confirmation bias from determining the conclusions of a
meta-analysis (Lakens et al., 2016). We report deviations from our
preregistered protocol at the end of the Method section. We make data,
analysis scripts, and the fully reproducible manuscript available
online.\footnote{\url{https://osf.io/w5tqv}}

@@

\textbf{\emph{Study Design.}} We categorized each study as either
observational and cross-sectional, observational and longitudinal,
quasi-experimental, experimental, an intervention, or other
(\(\kappa = .85\)).\footnote{While the term `observational study' is
  sometimes used to refer to the naturalistic observation of behavior,
  we use it in its broader sense, that is, to refer to studies that do
  not involve random assignment into experimental conditions (Cochran,
  1965; Rosenbaum, 2005).} We categorized each sample as either a
student convenience sample, a non-student convenience sample, a
probability or representative sample, or another kind of sample
(\(\kappa = .73\)). We categorized the age group(s) in each sample as
children (\(\leq 12\) years), adolescents (13--18 years), or adults
(\(\geq 18\) years; \(\kappa = .64\)).

@@

Figure 2 provides a qualitative overview of the literature. Even though
most participants came from studies conducted in India (\(N = 104,639\))
and the United States (\(N = 60,983\); Figure 2a), most samples were
collected in North America (\(I = 48\)) and Europe (\(I = 43\)), with
few samples from South America and Africa (Figure 2b). Most samples
focused on relative inequalities that resulted from long-term migration
(\(I = 59\)), slavery (\(I = 19\)), colonization (\(I = 19\)), and
short-term migration (\(I = 14\)). With few exceptions, studies used
observational, cross-sectional survey designs. Samples sizes ranged from
64 to 49,764 (\(\textit{Mdn} = 287\)). Unlike most psychological
research, 2 in 5 samples used probability or representative sampling. Of
all samples, 55\% (\(I = 77\)) were collected with the intention to
examine the effects of intergroup contact on the outcomes considered in
this meta-analysis. Many studies instead examined acculturation
processes (Berry, 1997) and measured intergroup contact and perceived
injustice without hypothesizing that the former would affect the latter.
Together, these observations highlight both limitations and strengths of
the empirical literature.

@@

Third, the available evidence is almost entirely from cross-sectional,
observational studies and thus consistent with alternative explanations
for the observed associations. Support for social change could reduce
intergroup contact rather than the other way around. For example,
disadvantaged-group members involved in collective action might avoid
forming friendships with advantaged-group members. Alternatively, the
observed associations could be spurious with both being caused by an
unobserved confounder. For example, a disadvantaged-group member's
socioeconomic status might both expose them to more advantaged-group
members and reduce their perception of injustice. Future research should
prioritize longitudinal studies to confirm the direction of the observed
associations and (field) experiments to rule out confounding and other
alternative explanations.\footnote{Published research includes only
  three longitudinal studies (Koschate et al., 2012; Reimer et al.,
  2017; Tropp et al., 2012), all of which conflated within-person change
  and between-person stability (Hamaker et al., 2015), two experimental
  studies with a no-contact control condition (Becker et al., 2013;
  Droogendyk et al., 2016), and two intervention studies (Reimer et al.,
  2021; Shani \& Boehnke, 2017).}

@@

Finally, we cannot rule out that biases in the literature caused our
meta-analysis to overestimate the `ironic' effects of intergroup
contact. Publication bias could have prevented studies that found
positive or non-significant associations from entering the published
literature. We used various methods to detect and correct for
publication bias, yet did not find conclusive evidence for or against
publication bias. That said, 61\% of samples were either not collected
with the intention to study the effects of intergroup contact on a
relevant outcome (45\%) or from unpublished studies (32\%)---and thus
unlikely to have been affected by publication bias. Another bias that
could have affected our findings is that researchers made data-dependent
decisions while collecting and analyzing data that are known to inflate
false-positive findings (Gelman \& Loken, 2014; Simmons et al., 2011).
This risk is pertinent to a literature in which most studies are
observational, include many measures, and are not preregistered---and,
therefore, do not constrain the direction of observed associations or
the selection of included variables. Future research should use
preregistration and other methods to prevent undisclosed flexibility in
data collection and analysis.

@@

@@

\leavevmode\hypertarget{ref-cochran_planning_1965}{}%
Cochran, W. G. (1965). The planning of observational studies of human
populations. \emph{Journal of the Royal Statistical Society},
\emph{128}(2), 234--266. \url{https://doi.org/10.2307/2344179}

@@

\leavevmode\hypertarget{ref-everitt_observational_2005}{}%
Rosenbaum, P. R. (2005). Observational study. In B. S. Everitt \& D. C.
Howell (Eds.), \emph{Encyclopedia of statistics in behavioral science}
(3rd ed., pp. 1451--1462). John Wiley \& Sons, Ltd.
\url{https://doi.org/10.1002/0470013192}

>>>>

We prepared a protocol in accordance with the PRISMA-P statement (Moher
et al., 2015) and preregistered it on the Open Science
Framework.\footnote{\url{https://osf.io/ryuev/}} Preregistration
precludes undisclosed flexibility in study selection, outcome selection,
and other decisions that influence effect size estimates---and thus
prevents confirmation bias from determining the conclusions of a
meta-analysis (Lakens et al., 2016). We report deviations from our
preregistered protocol at the end of the Method section. We make data,
analysis scripts, and the fully reproducible manuscript available
online.\footnote{\url{https://osf.io/w5tqv/}}

@@

\textbf{\emph{Study Design.}} We categorized each study as either
correlational and cross-sectional, correlational and longitudinal,
quasi-experimental, experimental, an intervention, or other
(\(\kappa = .85\)). We categorized each sample as either a student
convenience sample, a non-student convenience sample, a probability or
representative sample, or another kind of sample (\(\kappa = .73\)). We
categorized the age group(s) in each sample as children (\(\leq 12\)
years), adolescents (13--18 years), or adults (\(\geq 18\) years;
\(\kappa = .64\)).

@@

Figure 2 provides a qualitative overview of the literature. Even though
most participants came from studies conducted in India (\(N = 104,639\))
and the United States (\(N = 60,983\); Figure 2a), most samples were
collected in North America (\(I = 48\)) and Europe (\(I = 43\)), with
few samples from South America and Africa (Figure 2b). Most samples
focused on relative inequalities that resulted from long-term migration
(\(I = 59\)), slavery (\(I = 19\)), colonization (\(I = 19\)), and
short-term migration (\(I = 14\)). With few exceptions, studies used
correlational, cross-sectional survey designs. Samples sizes ranged from
64 to 49,764 (\(\textit{Mdn} = 287\)). Unlike most psychological
research, 2 in 5 samples used probability or representative sampling. Of
all samples, 55\% (\(I = 77\)) were collected with the intention to
examine the effects of intergroup contact on the outcomes considered in
this meta-analysis. Many studies instead examined acculturation
processes (Berry, 1997) and measured intergroup contact and perceived
injustice without hypothesizing that the former would affect the latter.
Together, these observations highlight both limitations and strengths of
the empirical literature.

@@

Third, the available evidence is almost entirely from cross-sectional,
correlational studies and thus consistent with alternative explanations
for the observed associations. Support for social change could reduce
intergroup contact rather than the other way around. For example,
disadvantaged-group members involved in collective action might avoid
forming friendships with advantaged-group members. Alternatively, the
observed associations could be spurious with both being caused by an
unobserved confounder. For example, a disadvantaged-group member's
socioeconomic status might both expose them to more advantaged-group
members and reduce their perception of injustice. Future research should
prioritize longitudinal studies to confirm the direction of the observed
associations and (field) experiments to rule out confounding and other
alternative explanations.\footnote{Published research includes only
  three longitudinal studies (Koschate et al., 2012; Reimer et al.,
  2017; Tropp et al., 2012), all of which conflated within-person change
  and between-person stability (Hamaker et al., 2015), two experimental
  studies with a no-contact control condition (Becker et al., 2013;
  Droogendyk et al., 2016), and two intervention studies (Reimer et al.,
  2021; Shani \& Boehnke, 2017).}

@@

Finally, we cannot rule out that biases in the literature caused our
meta-analysis to overestimate the `ironic' effects of intergroup
contact. Publication bias could have prevented studies that found
positive or non-significant associations from entering the published
literature. We used various methods to detect and correct for
publication bias, yet did not find conclusive evidence for or against
publication bias. That said, 61\% of samples were either not collected
with the intention to study the effects of intergroup contact on a
relevant outcome (45\%) or from unpublished studies (32\%)---and thus
unlikely to have been affected by publication bias. Another bias that
could have affected our findings is that researchers made data-dependent
decisions while collecting and analyzing data that are known to inflate
false-positive findings (Gelman \& Loken, 2014; Simmons et al., 2011).
This risk is pertinent to a literature in which most studies are
correlational, include many measures, and are not preregistered---and,
therefore, do not constrain the direction of observed associations or
the selection of included variables. Future research should use
preregistration and other methods to prevent undisclosed flexibility in
data collection and analysis.

@@

\leavevmode\hypertarget{ref-osf_repository}{}%
Blinded for Review. (2022). \emph{Meta-analysis of the {`ironic'}
effects of intergroup contact} {[}Dataset{]}.
\url{https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a}

@@

@@
